---
# NOTE: the YAML header is not used to create the title slide. The title slide 
# will be created manually as the first slide, and meta tags are specified 
# using the metathis package (see below)
title: "Doing more with data: An introduction to Arrow for R users"
output:
  xaringan::moon_reader:
    seal: false 
    lib_dir: libs
    css: ["style/xaringan-themer.css", "style/voltron-data-style.css"]
    nature:
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc((%current% - 1) / (%total% - 1)* 100%);">
          </div>
        </div>
---


```{r load-packages, message=FALSE, echo=FALSE}
library(xaringanthemer)
library(tidyverse)
library(flair)
library(metathis)
```


```{r meta, echo=FALSE}
# This is used to create all the social media tags etc. If you don't have a 
# good social preview image of your own, there's a generic one that you can use
# in this repo. Just substitute this:
#
#   image = "https://djnavarro.net/xaringan-template-voltrondata/img/social-media-preview.png",
#
# It won't show anything specific to your slides but it will show the Voltron 
# Data logo and data lake image against the usual dark green background

meta() %>%
  meta_name("github-repo" = "djnavarro/slides-data-thread-2022") %>% 
  meta_social(
    title = "Doing more with data: An introduction to Arrow for R users",
    description = "An example slide deck using the Voltron Data template for xaringan",
    url = "https://djnavarro.net/xaringan-template-voltrondata",
    image = "https://djnavarro.net/xaringan-template-voltrondata/img/title-slide-image.png",
    image_alt = "Image showing the title slide. The title in white reads 'Voltron Data template for xaringan users', with 'by Danielle Navarro' immediately underneath. The background is dark green, with the Voltron Data 'data lake' image shown in light green below the title, and the company logo shown in light green in the top right corner",
    og_type = "website",
    og_author = "Danielle Navarro",
    twitter_card_type = "summary_large_image",
    twitter_creator = "@djnavarro"
  )
```


```{r child = "style/slide-setup.Rmd"}
# Much of the set up work defining the theme is done in slide-setup.Rmd
```


class: title-slide, theme-green-maximal  <!-- title slide style -->


.center[.large[.boldface[Doing more with data: An introduction to Arrow for R users]]]  <!-- title, one line -->

.center[.midi[by Danielle Navarro]] <!-- author, one line -->
.center[.midi[[github.com/djnavarro/slides-data-thread-2022](https://github.com/djnavarro/slides-data-thread-2022)]] <!-- general info, one line -->

<!--
font-size classes (relative to normal 100%)

.larger = 200% 
.large  = 130% 
.midi   = 85% 
.small  = 70% 
.xsmall = 60% 
.tiny   = 50% 
-->


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## Who is Danielle Navarro?

- Developer advocate at Voltron Data
- Data scientist and generative artist
- General nuisance in the R community
- Tweets incessantly at [@djnavarro](https://twitter.com/djnavarro)



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## What is Apache Arrow?

- Toolbox for in-memory analytics
- Toolbox for larger than memory data
- Toolbox for connecting languages and hardware
- Efficient, performant, open source
- Details at [arrow.apache.org](https://arrow.apache.org)



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## Who is this talk for?

- R users who want to understand Apache Arrow
- Arrow developers who want to understand the R community
- Anyone who wants to see me make a fool of myself




<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## What does this talk cover?

- The "big idea" behind Arrow
- An overview of the {arrow} R package
- Example analyses using NYC taxi data set (~1.7 billion rows)
- Links to additional resources!




<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-light-maximal

## Apache Arrow: The Big Idea





<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-maximal

## class: theme-green-maximal

- good for section titles
- especially with helper classes added



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## class: theme-green

- good for content slides with very little text or code
- links display in orange, https://arrow.apache.org
- inline code is light green, `library(arrow)`



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-minimal

## class: theme-green-minimal

- good for content slides with more text or code
- links display in orange, https://arrow.apache.org
- inline code is light green, `library(arrow)`
- code blocks use solarized-dark theme:

```{r}
runif(1000) |>
  sqrt() |>
  quantile(probs = seq(0, 1, .1))
```


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-minimal

## class: theme-green-minimal

this is what block quotes look like:

> lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum




<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-light

## class: theme-light

- theme-light mirrors theme-green in structure
- there is also a theme-light-minimal and theme-light-maximal
- links display in darker orange, https://arrow.apache.org
- inline code is darker green, `library(arrow)`



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-light-minimal

## class: theme-light-minimal

.pull-left[
- pull-left makes a left column
- pull-right makes a right column
- code can nest within a column:
```{r}
sydney_time <- lubridate::now()
berlin_time <- sydney_time |> 
  lubridate::with_tz("Europe/Berlin")
```
]

--

.pull-right[
- handy to show code on the left
- then reveal results on the right:
```{r, results='hold'}
sydney_time
berlin_time
```
]


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-dark

## class: theme-dark

- theme-dark, theme-dark-minimal, and theme-dark-maximal
- links display in orange, https://arrow.apache.org
- inline code is light green, `library(arrow)`


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-dark

## class: theme-dark

.pull-left-wide[
- pull-left-wide makes a 70%-width column
- pull-right-wide makes a 70%-width column
- pull-left-narrow makes a 30%-width column
- pull-right-narrow makes a 30%-width column
]

--

.pull-right-narrow[
```{r}
runif(1000) |>
  sqrt() |>
  mean()
```
]


---

class: theme-dark-nologo

## class: theme-dark-minimal-nologo

- every theme has a "nologo" version
- prevents title and logo competing for audience attention


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-minimal

## A slide with flair

- the flair package allows fancy code decoration
- docs: https://r-for-educators.github.io/flair/

```{r filenames, include=FALSE, results='hide'}
"analysis01_descriptive-statistics.R"
"analysis02_preregistered-analysis.R"
"notes01_realising-the-problem.txt"
"analysis03_departing-from-the-plan.R"
"notes02_tentative-write-up.docx"
```

.pull-left-wide[
<!-- 
The pull left wide class is used here for aesthetic purposes. It 
prevents the background colour from the code block extending too 
far rightwards
-->
```{r, echo=FALSE, results='asis'}
decorate_chunk("filenames") %>% 
  flair_rx("_[^.]*", background = "#000000") %>% 
  flair::knit_print.with_flair()
```
]


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-minimal

## A slide with ggplot2 themes

.pull-left[
```{r define-ggplot-example, fig.show='hide'}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(
    x = "engine displacement", 
    y = "highway mileage",
    title = "An example plot",
    subtitle = "Theme based on theme_bw()"
  ) + 
  facet_wrap(vars(year)) + 
  theme_voltron_light()
```
]

--

.pull-right[
```{r run-ggplot-example, ref.label="define-ggplot-example", echo=FALSE}
# In case you are unfamiliar with it, this is the so-called "ref.label trick". 
# When knit, R will first execute code in the "ggplot-example" chunk and then
# any additional code in this chunk. It's especially handy in xaringan slides
# when you want to separate the code defining a plot from its execution. It's
# what allows me to put the code inside the .pull-left class, and the output
# in the .pull-right class
```
]



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-light-minimal

## A slide with ggplot2 themes

.pull-left[
```{r define-ggplot-example-light, fig.show='hide'}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(
    x = "engine displacement", 
    y = "highway mileage",
    title = "An example plot",
    subtitle = "Theme based on theme_dark()"
  ) + 
  facet_wrap(vars(year)) + 
  theme_voltron_dark()
```
]

.pull-right[
```{r run-ggplot-example-light, ref.label="define-ggplot-example-light", echo=FALSE}
```
]



<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-dark-minimal

## A slide with ggplot2 themes

.pull-left[
```{r define-ggplot-example-dark, fig.show='hide'}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(
    x = "engine displacement", 
    y = "highway mileage",
    title = "An example plot",
    subtitle = "Theme based on theme_dark()"
  ) + 
  facet_wrap(vars(year)) + 
  theme_voltron_light()
```
]

.pull-right[
```{r run-ggplot-example-dark, ref.label="define-ggplot-example-dark", echo=FALSE}
```
]


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green

## Using font-awesome icons in lists

<!-- 
embedding raw html is handy for things like using font-awesome icons as
the bullets in list 
-->
<ol class="fa-ul">
  <li><span class="fa-li"><i class="fas fa-check-square"></i></span>Completed task</li>
  <li><span class="fa-li"><i class="fas fa-check-square"></i></span>Completed task</li>
  <li><span class="fa-li"><i class="fas fa-spinner fa-pulse"></i></span>Work in progress</li>
  <li><span class="fa-li"><i class="far fa-square"></i></span>Future work</li>
</ol>


<!----------------------------- SLIDE BREAK ----------------------------->
---

class: theme-green-maximal

## Final slide with key points and links

.pull-left[
- xaringan templates are cool
- this is a work in progress
- please send bug reports!
]

.pull-right[
<ol class="fa-ul">
  <li><span class="fa-li"><i class="fa fa-twitter"></i></span>twitter.com/djnavarro</li>
  <li><span class="fa-li"><i class="fa fa-github"></i></span>github.com/djnavarro</li>
  <li><span class="fa-li"><i class="fa fa-globe"></i></span>djnavarro.net</li>
</ol>
]





